---
title: "MA 331 Final"
author: "Max Surets, Kera McGovern, and Adrian Gomes"
date: "12/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
```


```{r}
setwd(getwd())

library(readxl)
loss <- read_excel("./ex11-53cheese.xls")

taste = loss$taste
acetic = loss$acetic
h2s = loss$h2s
lactic = loss$lactic

```

# 11.53

### Taste  
Mean = `r mean(taste)`  
Median = `r median(taste)`  
Standard deviation = `r sd(taste)`  
Interquartile range = `r IQR(taste)`  

```{r out.width="50%"}
stem(taste)
qqnorm(taste, pch = 1, frame = FALSE)
qqline(taste, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line and the leaf plot looks skewed right.  

### Acetic
Mean = `r mean(acetic)`   
Median = `r median(acetic)`  
Standard deviation = `r sd(acetic)`  
Interquartile range = `r IQR(acetic)`  
```{r out.width = "50%"}
stem(acetic)
qqnorm(acetic, pch = 1, frame = FALSE)
qqline(acetic, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot does not look normal.


### H2S
Mean = `r mean(h2s)`   
Median = `r median(h2s)`  
Standard deviation = `r sd(h2s)`  
Interquartile range = `r IQR(h2s)`  
```{r out.width = "50%"}
stem(h2s)
qqnorm(h2s, pch = 1, frame = FALSE)
qqline(h2s, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot is skewed right. 



### Lactic
Mean = `r mean(lactic)`   
Median = `r median(lactic)`  
Standard deviation = `r sd(lactic)`  
Interquartile range = `r IQR(lactic)`  
```{r out.width = "50%"}
stem(lactic)
qqnorm(lactic, pch = 1, frame = FALSE)
qqline(lactic, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot is approximately normal.

In summary, taste, acetic, and H2S do not seem to be normally distributed, but lactic does.  





# 11.54

```{r out.width = "50%"}
plot(taste, acetic)
res = summary(aov(acetic~taste))
```

The relationship between acetic and taste is positive.  
Correlation = `r cor(taste, acetic)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(taste, h2s)
res = summary(aov(h2s~taste))
```

The relationship between h2s and taste is positive.  
Correlation = `r cor(taste, h2s)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(taste, lactic)
res = summary(aov(lactic~taste))
```

The relationship between lactic and taste is positive.  
Correlation = `r cor(taste, lactic)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(acetic, h2s)
res = summary(aov(acetic~h2s))
```

The relationship between acetic and h2s is positive.  
Correlation = `r cor(acetic, h2s)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(acetic, lactic)
res = summary(aov(acetic~lactic))
```

The relationship between acetic and lactic is positive.  
Correlation = `r cor(acetic, lactic)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(h2s, lactic)
res = summary(aov(h2s~lactic))
```

The relationship between h2s and lactic is positive.  
Correlation = `r cor(h2s, lactic)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  



# 11.55
$$H_0: \beta_1 = 0$$
$$H_A: \beta_1 \not= 0$$
```{r}
fstats <- c()
pvalues <- c()
rsq <- c()
s <- c()

tamodel <- lm(taste ~ acetic, data = loss)
modelSummary <- summary(tamodel)
fstats <- append(fstats, modelSummary$fstatistic[[1]])
s <- append(s, sqrt(modelSummary$sigma))
rsq <- append(rsq, modelSummary$r.squared)
modelCoeffs <- modelSummary$coefficients
t_value <- modelCoeffs["acetic", "t value"]
p_value <- modelSummary[["coefficients"]][8]
pvalues <- append(pvalues, p_value)
```
$\hat Taste$ = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["acetic"]` * Acetic  
T-value: `r t_value`  
P-value: `r p_value` 
  
Since the P-value of `r p_value` is less than the significance level of 0.05, the null hypothesis can be rejected as there is statistically significant evidence to suggest that there exists a positive relationship between acetic and taste.  

```{r out.width = "50%"}
Taste_Acetic <- lm(taste ~ acetic, data = loss)
modelSummary <- summary(Taste_Acetic)
hist(Taste_Acetic$residuals, pch = 16, col = "red", main = "A: Histogram of Residuals")

predictedy <- Taste_Acetic$fitted.values
plot(acetic,taste, main = "B: Least Squares of Taste v. Acetic")
abline(Taste_Acetic$coefficients["(Intercept)"], Taste_Acetic$coefficients["acetic"])

plot(Taste_Acetic$residuals, h2s, main = "C: Residuals v. H2S")
plot(Taste_Acetic$residuals, lactic, main = "D: Residuals v. Lactic")
```

As shown in Graph A, the residuals have an approximately Normal distribution. As seen in Graphs C and D, there is a positive association between the residuals of taste and acetic and the values of H2S and lactic, respectively.





# 11.56

$$H_0: \beta_1 = 0$$
$$H_A: \beta_1 \not= 0$$
```{r}
tamodel <- lm(taste ~ h2s, data = loss)
modelSummary <- summary(tamodel)
fstats <- append(fstats, modelSummary$fstatistic[[1]])
rsq <- append(rsq, modelSummary$r.squared)
s <- append(s, sqrt(modelSummary$sigma))
modelCoeffs <- modelSummary$coefficients
t_value_h2s <- modelCoeffs["h2s", "t value"]
p_value_h2s <- modelSummary[["coefficients"]][8]
pvalues <- append(pvalues, p_value_h2s)
```
$\hat Taste$ = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["h2s"]` * H2S  
T-value: `r t_value_h2s`  
P-value: `r p_value_h2s` 
  
Since the P-value of `r p_value_h2s` is less than the significance level of 0.05, the null hypothesis can be rejected as there is statistically significant evidence to suggest that there exists a positive relationship between H2S and taste.  

```{r out.width = "50%"}
Taste_H2s <- lm(taste ~ h2s, data = loss)
modelSummary <- summary(Taste_H2s)
hist(Taste_H2s$residuals, pch = 16, col = "red", main = "A: Histogram of Residuals")

predictedy <- Taste_H2s$fitted.values
plot(h2s,taste, main = "B: Least Squares of Taste v. H2S")
abline(Taste_H2s$coefficients["(Intercept)"], Taste_H2s$coefficients["h2s"])

plot(Taste_H2s$residuals, acetic, main = "C: Residuals v. Acetic")
plot(Taste_H2s$residuals, lactic, main = "D: Residuals v. Lactic")
```

As shown in Graph A, the residuals are approximately Normally distributed. As seen in Graphs C and D, there appears to be an extremely weak positive correlation between the residuals of taste and H2s and the values of acetic and lactic, respectively.




# 11.57

$$H_0: \beta_1 = 0$$
$$H_A: \beta_1 \not= 0$$
```{r}
tamodel <- lm(taste ~ lactic, data = loss)
modelSummary <- summary(tamodel)
fstats <- append(fstats, modelSummary$fstatistic[[1]])
rsq <- append(rsq, modelSummary$r.squared)
s <- append(s, sqrt(modelSummary$sigma))
modelCoeffs <- modelSummary$coefficients
t_value_lactic <- modelCoeffs["lactic", "t value"]
p_value_lactic <- modelSummary[["coefficients"]][8]
pvalues <- append(pvalues, p_value_lactic)
```
$\hat Taste$ = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["lactic"]` * Lactic  
T-value: `r t_value_lactic`  
P-value: `r p_value_lactic` 
  
Since the P-value of `r p_value_lactic` is less than the significance level of 0.05, the null hypothesis can be rejected as there is statistically significant evidence to suggest that there exists a positive relationship between lactic and taste.  

```{r out.width = "50%"}
Taste_L <- lm(taste ~ lactic, data = loss)
modelSummary <- summary(Taste_L)
hist(Taste_L$residuals, pch = 16, col = "red", main = "A: Histogram of Residuals")

predictedy <- Taste_L$fitted.values
plot(lactic,taste, main = "B: Least Squares of Taste v. Lactic")
abline(Taste_L$coefficients["(Intercept)"], Taste_L$coefficients["lactic"])

plot(Taste_L$residuals, acetic, main = "C: Residuals v. Acetic")
plot(Taste_L$residuals, h2s, main = "D: Residuals v. H2S")
```

As shown in Graph A, the residuals are approximately Normally distributed. As seen in Graphs C and D, there appears to no correlation between the residuals of taste and H2s and the values of acetic and H2S, respectively.

\   
\   

# 11.58
```{r out.width = "50%"}
compTable = data.frame("F statistic"=fstats, "P-value"=pvalues, "R-squared"=rsq, "s"=s)
knitr::kable(compTable)
```

$\hat Taste=-61.4986124+15.6477672*Acetic$  
$\hat Taste=-9.7868374+5.7760886*H2S$  
$\hat Taste=-29.8588324+37.7199485*Lactic$  

The intercepts in these three equations are different because of differences in the variations of the response variable. For example, the values of H2S at low values of taste were concentrated close to the line compared to the values of the response variable acetic, which were scattered far from the line. 

# 11.59
$$H_0: \beta_1 = \beta_2 = 0$$
$$H_A: \beta_j \not=0\mid j \in \{1,2\}$$
```{r}
tamodel <- lm(taste ~ acetic + h2s, data = loss)
modelSummary <- summary(tamodel)
modelCoeffs <- modelSummary$coefficients

beta <- modelCoeffs["acetic", "Estimate"]
beta2 <- modelCoeffs["h2s", "Estimate"] 

se <- modelCoeffs["acetic", "Std. Error"]
se2 <- modelCoeffs["h2s", "Std. Error"]

t_value <- beta/se #what to do with second values?
p_value <- 2*pt(-abs(t_value), df=nrow(loss) - 4)
```
Taste = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["acetic"]` * Acetic + `r tamodel$coefficients["h2s"]` * H2S  
$R^2=$ `r modelSummary$r.squared`  
Acetic T-value: `r t_value`  
Acetic P-value: `r p_value`  

In the model where Acetic is the only predictor, the significance of Acetic is 0.001658, which is lower than the significance of Acetic in the model where Acetic and H2S are both used, where the it is 0.4065277. The model where both are used is stronger because it has a much lower P-value ($7.645*10^{-6}$) and the $R^2$ value is higher (`r modelSummary$r.squared`). Acetic appears to be a good predictor alone but not with H2S because the variables are strongly correlated (r = 0.6179559). Due to this, acetic does not provide much significant additional information.


# 11.60

$$H_0: \beta_1 = \beta_2 = 0$$
$$H_A: \beta_j \not=0\mid j \in \{1,2\}$$

```{r}
tamodel <- lm(taste ~ h2s + lactic, data = loss)
modelSummary <- summary(tamodel)
modelCoeffs <- modelSummary$coefficients

t_value <- modelCoeffs["h2s", "t value"]
t_value2 <- modelCoeffs["lactic", "t value"] 

p_value <- modelCoeffs[[11]]
p_value2 <- modelCoeffs[[12]]

f <- summary(tamodel)$fstatistic
p_value_model <- pf(f[1],f[2],f[3],lower.tail=F)
```
Taste = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["h2s"]` * H2S + `r tamodel$coefficients["lactic"]` * Lactic  
$R^2=$ `r modelSummary$r.squared`  

H2S T-value: `r t_value`  
H2S P-value: `r p_value`  
Lactic T-value: `r t_value2`  
Lactic P-value: `r p_value2`  
F-statistic P-value: `r p_value_model`  

Using both H2S and Lactic as predictors in the model resulted in a better model than that of each of the explanatory variables on their own. As seen above, the p-value obtained when H2S and lactic are both used as predictors is `r p_value_model` When they used as individual explanatory variables, their p-values are `r p_value_h2s` and `r p_value_lactic` for H2S and lactic, respectively. When they are both used as predictors together, the p-value is lower which shows that the model is more significant and thus more reliable as a predictor for the taste of the cheese.



# 11.61
$$H_0: \beta_1 = \beta_2 = \beta_3 = 0$$
$$H_A: \beta_j \not=0\mid j \in \{1,2,3\}$$
```{r}
tamodel <- lm(taste ~ acetic + h2s + lactic, data = loss)
modelSummary <- summary(tamodel)
modelCoeffs <- modelSummary$coefficients

f <- summary(tamodel)$fstatistic
p_value_model <- pf(f[1],f[2],f[3],lower.tail=F)
```

Taste = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["acetic"]` * Acetic + `r tamodel$coefficients["h2s"]` * H2S + `r tamodel$coefficients["lactic"]` * Lactic  

Acetic T-value = `r modelCoeffs[10]`  
H2S T-value = `r modelCoeffs[11]`  
Lactic T-value = `r modelCoeffs[12]`  

Acetic P-value = `r modelCoeffs[14]`  
H2S P-value = `r modelCoeffs[15]`  
Lactic P-value = `r modelCoeffs[16]`  


The model has an R-squared value of `r modelSummary$r.squared`. There is a strong positive relationship between the explanatory variables and taste. The model has a P-value of `r p_value_model`. At the significance level of 0.05, we have statistically significant evidence to suggest that a positive relationship exists between the taste and explanatory variables acetic, H2S, and lactic.  


```{r out.width="50%"}
plot(tamodel, which=2)
```

The Q-Q plot shows that the distribution of the residuals is approximately normal.

```{r out.width="50%"}
plot(tamodel, which=1)
```

The Residuals vs Fitted graph shows a relatively uniform distribution with no obvious pattern. The distribution of the residuals can be described as approximately normal.  
\   

Since the P-value for acetic was so high (`r modelCoeffs[14]`), we should consider removing it from the model in order to increase precision.  

From all of the previous linear models, the model with H2S and Lactic as explanatory variables for Taste should be preferred. It is the most significant model because it has the lowest P-value ($6.5513712*10^{-7}$).


