---
title: "MA 331 Final"
author: "Max Surets, Kera McGovern, and Adrian Gomes"
date: "12/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
```


```{r}
setwd(getwd())

library(readxl)
loss <- read_excel("./ex11-53cheese.xls")

taste = loss$taste
acetic = loss$acetic
h2s = loss$h2s
lactic = loss$lactic

```

# 11.53

### Taste  
Mean = `r mean(taste)`  
Median = `r median(taste)`  
Standard deviation = `r sd(taste)`  
Interquartile range = `r IQR(taste)`  

```{r out.width="50%"}
stem(taste)
qqnorm(taste, pch = 1, frame = FALSE)
qqline(taste, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line and the leaf plot looks skewed right.  

### Acetic
Mean = `r mean(acetic)`   
Median = `r median(acetic)`  
Standard deviation = `r sd(acetic)`  
Interquartile range = `r IQR(acetic)`  
```{r out.width = "50%"}
stem(acetic)
qqnorm(acetic, pch = 1, frame = FALSE)
qqline(acetic, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot does not look normal.


### H2S
Mean = `r mean(h2s)`   
Median = `r median(h2s)`  
Standard deviation = `r sd(h2s)`  
Interquartile range = `r IQR(h2s)`  
```{r out.width = "50%"}
stem(h2s)
qqnorm(h2s, pch = 1, frame = FALSE)
qqline(h2s, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot is skewed right. 



### Lactic
Mean = `r mean(lactic)`   
Median = `r median(lactic)`  
Standard deviation = `r sd(lactic)`  
Interquartile range = `r IQR(lactic)`  
```{r out.width = "50%"}
stem(lactic)
qqnorm(lactic, pch = 1, frame = FALSE)
qqline(lactic, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot is approximately normal.

In summary, taste, acetic, and H2S do not seem to be normally distributed, but lactic does.  





# 11.54

```{r out.width = "50%"}
plot(taste, acetic)
res = summary(aov(taste~acetic))
```

The relationship between taste and acetic is positive.  
Correlation = `r cor(taste, acetic)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(taste, h2s)
res = summary(aov(taste~h2s))
```

The relationship between taste and h2s is positive.  
Correlation = `r cor(taste, h2s)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(taste, lactic)
res = summary(aov(taste~lactic))
```

The relationship between taste and lactic is positive.  
Correlation = `r cor(taste, lactic)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(acetic, h2s)
res = summary(aov(acetic~h2s))
```

The relationship between acetic and h2s is positive.  
Correlation = `r cor(acetic, h2s)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(acetic, lactic)
res = summary(aov(acetic~lactic))
```

The relationship between acetic and lactic is positive.  
Correlation = `r cor(acetic, lactic)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  

```{r out.width = "50%"}
plot(h2s, lactic)
res = summary(aov(h2s~lactic))
```

The relationship between h2s and lactic is positive.  
Correlation = `r cor(h2s, lactic)`  
P-value = `r res[[1]][["Pr(>F)"]][1]`  



# 11.55
Assume a significance level of 0.05:
$$H_0: \beta_1 = 0$$
$$H_A: \beta_1 \not= 0$$
```{r}
tamodel <- lm(taste ~ acetic, data = loss)
modelSummary <- summary(tamodel)
modelCoeffs <- modelSummary$coefficients
beta <- modelCoeffs["acetic", "Estimate"]
se <- modelCoeffs["acetic", "Std. Error"]
t_value <- beta/se
p_value <- 2*pt(-abs(t_value), df=nrow(loss) - 3)
```
Taste = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["acetic"]` * Acetic  
T-value: `r t_value`  
P-value: `r p_value` 
  
Since the p value of `r p_value` is less than the significance level of 0.05, the null hypothesis can be rejected as the test is statistically significant.  

```{r out.width = "50%"}
Taste_Acetic <- lm(taste ~ acetic, data = loss)
modelSummary <- summary(Taste_Acetic)
hist(Taste_Acetic$residuals, pch = 16, col = "red", main = "A: Histogram of Residuals")

predictedy <- Taste_Acetic$fitted.values
plot(acetic,taste, main = "B: Least Squares of Taste v. Acetic")
abline(Taste_Acetic$coefficients["(Intercept)"], Taste_Acetic$coefficients["acetic"])

plot(Taste_Acetic$residuals, h2s, main = "C: Residuals v. H2S")
plot(Taste_Acetic$residuals, lactic, main = "D: Residuals v. Lactic")
```

As shown in Graph A, the residuals have an approximately Normal distribution. As seen in Graphs C and D, there is a positive association between the residuals of taste and acetic and the values of H2S and lactic, respectively.





# 11.56

Assume a significance level of 0.05:
$$H_0: \beta_1 = 0$$
$$H_A: \beta_1 \not= 0$$
```{r}
tamodel <- lm(taste ~ h2s, data = loss)
modelSummary <- summary(tamodel)
modelCoeffs <- modelSummary$coefficients
beta <- modelCoeffs["h2s", "Estimate"]
se <- modelCoeffs["h2s", "Std. Error"]
t_value <- beta/se
p_value <- 2*pt(-abs(t_value), df=nrow(loss) - 3)
```
Taste = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["h2s"]` * H2S  
T-value: `r t_value`  
P-value: `r p_value` 
  
Since the p value of `r p_value` is less than the significance level of 0.05, the null hypothesis can be rejected as the test is statistically significant.  

```{r out.width = "50%"}
Taste_H2s <- lm(taste ~ h2s, data = loss)
modelSummary <- summary(Taste_H2s)
hist(Taste_H2s$residuals, pch = 16, col = "red", main = "A: Histogram of Residuals")

predictedy <- Taste_H2s$fitted.values
plot(h2s,taste, main = "B: Least Squares of Taste v. H2S")
abline(Taste_H2s$coefficients["(Intercept)"], Taste_H2s$coefficients["h2s"])

plot(Taste_H2s$residuals, acetic, main = "C: Residuals v. Acetic")
plot(Taste_H2s$residuals, lactic, main = "D: Residuals v. Lactic")
```

As shown in Graph A, the residuals are approximately Normally distributed. As seen in Graphs C and D, there appears to be an extremely weak positive correlation between the residuals of taste and H2s and the values of acetic and lactic, respectively.




# 11.57

Assume a significance level of 0.05:
$$H_0: \beta_1 = 0$$
$$H_A: \beta_1 \not= 0$$
```{r}
tamodel <- lm(taste ~ lactic, data = loss)
modelSummary <- summary(tamodel)
modelCoeffs <- modelSummary$coefficients
beta <- modelCoeffs["lactic", "Estimate"]
se <- modelCoeffs["lactic", "Std. Error"]
t_value <- beta/se
p_value <- 2*pt(-abs(t_value), df=nrow(loss) - 3)
```
Taste = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["lactic"]` * Lactic  
T-value: `r t_value`  
P-value: `r p_value` 
  
Since the p value of `r p_value` is less than the significance level of 0.05, the null hypothesis can be rejected as the test is statistically significant.  

```{r out.width = "50%"}
Taste_L <- lm(taste ~ lactic, data = loss)
modelSummary <- summary(Taste_L)
hist(Taste_L$residuals, pch = 16, col = "red", main = "A: Histogram of Residuals")

predictedy <- Taste_L$fitted.values
plot(lactic,taste, main = "B: Least Squares of Taste v. Lactic")
abline(Taste_L$coefficients["(Intercept)"], Taste_L$coefficients["lactic"])

plot(Taste_L$residuals, acetic, main = "C: Residuals v. Acetic")
plot(Taste_L$residuals, h2s, main = "D: Residuals v. H2S")
```

As shown in Graph A, the residuals are approximately Normally distributed. As seen in Graphs C and D, there appears to no correlation between the residuals of taste and H2s and the values of acetic and H2S, respectively.






# 11.58



# 11.59
Assume a significance level of 0.05:
$$H_0: \beta_1 = \beta_2 = 0$$
$$H_A: \beta_1 \not= \beta_2 \not= 0$$
```{r}
tamodel <- lm(taste ~ acetic + h2s, data = loss)
modelSummary <- summary(tamodel)
summary(tamodel)
modelCoeffs <- modelSummary$coefficients

beta <- modelCoeffs["acetic", "Estimate"]
beta2 <- modelCoeffs["h2s", "Estimate"] 

se <- modelCoeffs["acetic", "Std. Error"]
se2 <- modelCoeffs["h2s", "Std. Error"]

t_value <- beta/se #what to do with second values?
p_value <- 2*pt(-abs(t_value), df=nrow(loss) - 4)
```
Taste = `r tamodel$coefficients["(Intercept)"]` + `r tamodel$coefficients["acetic"]` * Acetic + `r tamodel$coefficients["h2s"]` * H2S  
T-value: `r t_value`  
P-value: `r p_value`


# 11.60
# 11.61




