---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
```


```{r}
setwd(getwd())

library(readxl)
loss <- read_excel("./ex11-53cheese.xls")

taste = c()
acetic = c()
h2s = c()
lactic = c()

for (n in 1:nrow(loss)) {
  taste <- append(taste, as.numeric(loss[n,1]))
  acetic <- append(acetic, as.numeric(loss[n,2]))
  h2s <- append(h2s, as.numeric(loss[n,3]))
  lactic <- append(lactic, as.numeric(loss[n,4]))
}


```

# 11.53

### Taste  
Mean = `r mean(taste)`  
Median = `r median(taste)`  
Standard deviation = `r sd(taste)` 
Interquartile range = `r IQR(taste)`
```{r out.width="50%"}
stem(taste)
qqnorm(taste, pch = 1, frame = FALSE)
qqline(taste, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line and the leaf plot looks approximately normal.  

### Acetic
Mean = `r mean(acetic)`   
Median = `r median(acetic)`  
Standard deviation = `r sd(acetic)`  
Interquartile range = `r IQR(acetic)`  
```{r out.width = "50%"}
stem(acetic)
qqnorm(acetic, pch = 1, frame = FALSE)
qqline(acetic, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot shows a right skew.


### H2S
Mean = `r mean(h2s)`   
Median = `r median(h2s)`  
Standard deviation = `r sd(h2s)`  
Interquartile range = `r IQR(h2s)`  
```{r out.width = "50%"}
stem(h2s)
qqnorm(h2s, pch = 1, frame = FALSE)
qqline(h2s, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot does not look to be normally distributed.



### Lactic
Mean = `r mean(lactic)`   
Median = `r median(lactic)`  
Standard deviation = `r sd(lactic)`  
Interquartile range = `r IQR(lactic)`  
```{r out.width = "50%"}
stem(lactic)
qqnorm(lactic, pch = 1, frame = FALSE)
qqline(lactic, col = "steelblue", lwd = 2)
```

The normal QQ plot shows a roughly straight line but the leaf plot shows a right skew.  

In summary, taste seems to be approximately normally distributed, but acetic, H2S, and lactic do not seem to be.  





# 11.54

```{r out.width = "50%"}
plot(taste, acetic)
```


```{r out.width = "50%"}
plot(taste, h2s)
```


```{r out.width = "50%"}
plot(taste, lactic)
```


```{r out.width = "50%"}
plot(acetic, h2s)

```


```{r out.width = "50%"}
plot(acetic, lactic)
```


```{r out.width = "50%"}
plot(h2s, lactic)
```








# 11.55
# 11.56
# 11.57
# 11.58
# 11.59
# 11.60
# 11.61




