---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
```


```{r}
setwd(getwd())

library(readxl)
loss <- read_excel("./ex11-53cheese.xls")

taste = c()
acetic = c()
h2s = c()
lactic = c()

for (n in 1:nrow(loss)) {
  taste <- append(taste, as.numeric(loss[n,1]))
  acetic <- append(acetic, as.numeric(loss[n,2]))
  h2s <- append(h2s, as.numeric(loss[n,3]))
  lactic <- append(lactic, as.numeric(loss[n,4]))
}


```

# 11.53

### Taste  
Mean = `r mean(taste)`  
Median = `r median(taste)`  
Standard deviation = `r sd(taste)` 
Interquartile range = `r IQR(taste)`
```{r out.width="50%"}
stem(taste)
qqnorm(taste, pch = 1, frame = FALSE)
qqline(taste, col = "steelblue", lwd = 2)
```


### Acetic
Mean = `r mean(acetic)`   
Median = `r median(acetic)`  
Standard deviation = `r sd(acetic)`  
Interquartile range = `r IQR(acetic)`  
```{r out.width = "50%"}
stem(acetic)
qqnorm(acetic, pch = 1, frame = FALSE)
qqline(acetic, col = "steelblue", lwd = 2)
```

### H2S
Mean = `r mean(h2s)`   
Median = `r median(h2s)`  
Standard deviation = `r sd(h2s)`  
Interquartile range = `r IQR(h2s)`  
```{r out.width = "50%"}
stem(h2s)
qqnorm(h2s, pch = 1, frame = FALSE)
qqline(h2s, col = "steelblue", lwd = 2)
```

### Lactic
Mean = `r mean(lactic)`   
Median = `r median(lactic)`  
Standard deviation = `r sd(lactic)`  
Interquartile range = `r IQR(lactic)`  
```{r out.width = "50%"}
stem(lactic)
qqnorm(lactic, pch = 1, frame = FALSE)
qqline(lactic, col = "steelblue", lwd = 2)
```



